Forcing for Isabelle

### 2018-04-12 blueprint.tex
aparentemente está concluida la prueba de Separación

  Now we calculate $\val(\tau,F)$.
  \begin{multline*}
    \val(\tau,G) = \\
    \val\bigl(\{x\in\dom(\pi)\times\PP : \exists \th,p : x =\lb\th,p\rb
    \y  \\
    \forall F\ \PP\text{-generic}, p\in F \limp
    M[F],[\val(\th,F),\val(\pi,F),\val(\sig,F)]\models
    \dB{0}\in \dB{1} \y\phi(\dB{0},\dB{1},\dB{2})\}\bigr).
  \end{multline*}
  Hence
  \begin{multline*}
    \val(\tau,G) = 
    \{\val(\th,G): \th\in\dom(\pi) \y  p \in G 
    \y \forall F\ \PP\text{-generic}, \\
   p\in F \longrightarrow
  M[F],[\val(\th,F),\val(\pi,F),\val(\sig,F)]\models
  \dB{0}\in \dB{1} \y\phi(\dB{0},\dB{1},\dB{2})\}.
  \end{multline*}
  This is included in
  \begin{multline*}
    \{\val(\th,G): \th\in\dom(\pi) \y  p \in G 
    \y  \\
   p\in G \longrightarrow
   M[G],[\val(\th,G),\val(\pi,G),\val(\sig,G)]\models
   \dB{0}\in \dB{1} \y\phi(\dB{0},\dB{1},\dB{2})\},
  \end{multline*}
  and in turn,  replacing $\val(\pi,F),\val(\sig,F)$ by their values,
  this is included  in 
  \[ 
    \{\val(\th,G): \th\in\dom(\pi)
    \y 
   M[G],[\val(\th,G),c,w]\models
   \dB{0}\in \dB{1} \y\phi(\dB{0},\dB{1},\dB{2})\}.
  \]
  By absoluteness, we have that $\val(\tau,G)$ is
  \[
    \{\val(\th,G): \th\in\dom(\pi)
    \y \val(\th,G)\in c \y 
   M[G],[\val(\th,G),c,w]\models
   \phi(\dB{0},\dB{1},\dB{2})\}.
  \]
  which equals 
  \[
    \{x: x\in c \y 
    M[G],[x,c,w]\models
   \phi(\dB{0},\dB{1},\dB{2})\}.
  \]
  since for every $x\in c=\val(\pi,G)$ there must
  exist some $\th\in\dom(\pi)$ with $\val(\th,G)=x$. Therefore
  $\val(\tau,G)\subseteq S$.
\end{proof}

>>>>

  Now we calculate $\val(\tau,F)$.
  \begin{multline*}
    \val(\tau,G) = \\
    \val\bigl(\{x\in\dom(\pi)\times\PP : \exists \th,p : x =\lb\th,p\rb
    \y  \\
    \forall F\ \PP\text{-generic}, p\in F \limp
    M[F],[\val(\th,F),\val(\pi,F),\val(\sig,F)]\models
    \dB{0}\in \dB{1} \y\phi(\dB{0},\dB{1},\dB{2})\}\bigr).
  \end{multline*}
  Hence
  \begin{multline}\label{eq:val_tau}
    \val(\tau,G) = 
    \{\val(\th,G): \th\in\dom(\pi) \y  p \in G 
    \y \forall F\ \PP\text{-generic}, \\
   p\in F \longrightarrow
  M[F],[\val(\th,F),\val(\pi,F),\val(\sig,F)]\models
  \dB{0}\in \dB{1} \y\phi(\dB{0},\dB{1},\dB{2})\}.
  \end{multline}
  This is included in
  \begin{multline*}
    \{\val(\th,G): \th\in\dom(\pi) \y  p \in G 
    \y  \\
   p\in G \longrightarrow
   M[G],[\val(\th,G),\val(\pi,G),\val(\sig,G)]\models
   \dB{0}\in \dB{1} \y\phi(\dB{0},\dB{1},\dB{2})\},
  \end{multline*}
  and in turn,  replacing $\val(\pi,F),\val(\sig,F)$ by their values,
  this is included  in 
  \[ 
    \{\val(\th,G): \th\in\dom(\pi)
    \y 
   M[G],[\val(\th,G),c,w]\models
   \dB{0}\in \dB{1} \y\phi(\dB{0},\dB{1},\dB{2})\}.
  \]
  By absoluteness, we have that $\val(\tau,G)$ is
  \[
    \{\val(\th,G): \th\in\dom(\pi)
    \y \val(\th,G)\in c \y 
   M[G],[\val(\th,G),c,w]\models
   \phi(\dB{0},\dB{1},\dB{2})\}.
  \]
  which equals 
  \[
    \{x: x\in c \y 
    M[G],[x,c,w]\models
   \phi(\dB{0},\dB{1},\dB{2})\}.
  \]
  since for every $x\in c=\val(\pi,G)$ there must
  exist some $\th\in\dom(\pi)$ with $\val(\th,G)=x$. Therefore
  $\val(\tau,G)\subseteq S$.

  Now let $x\in S$. Hence there exists $\lb\th,q\rb\in\pi$ such that
  $q\in G$ and $x=\val(\th,G)$. By the second version of the Truth
  Lemma, since 
  \[
  M[G],[\val(\th,G),\val(\pi,G),\val(\sig,G)]\models
  \dB{0}\in\dB{1}\y\phi(\dB{0},\dB{1},\dB{2}),
  \]
  there must exists $r\in G$ such that 
  \[
  M,[\PP,\leq,\1,r,\th,\pi,\sig]\models
  \forceisa(\dB{0},\dB{1},\dB{2},\dB{3},
  (\dB{4}\in\dB{5}\y\phi(\dB{4},\dB{5},\dB{6}))).
  \]
  Since $G$ is a filter, there is $p\in G$ such that $p\leq q, r$.
  By Lemma~\ref{lem:strengthen}, we have
  \[
  M,[\PP,\leq,\1,p,\th,\pi,\sig]\models
  \forceisa(\dB{0},\dB{1},\dB{2},\dB{3},
  (\dB{4}\in\dB{5}\y \phi(\dB{4},\dB{5},\dB{6}))),
  \]
  which by the Truth Lemma gives us
  \[
  \forall F\ \PP\text{-generic}, \\
   p\in F \longrightarrow
  M[F],[\val(\th,F),\val(\pi,F),\val(\sig,F)]\models
  \dB{0}\in \dB{1} \y\phi(\dB{0},\dB{1},\dB{2}).
  \]
  Using now (\ref{eq:val_tau}), this shows that $S\sbq \val(\tau,G)$,
  since $\th\in\dom(\pi)$. 
\end{proof}
